SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2019-04-17 design_impl.tex
Paragraph spacing

\subsection{Changes Made to SPP}
\label{Impl-Changes}
To modify the SPP design to suit our scheme, the following changes
were made:

\noindent \textbf{Exporting Features from SPP:} PPF uses the metadata 
specific to SPP, to build some of the perceptron features. These include
the lookahead depth, signature and the confidence counter. These features 
were made visible to PPF\\
%
\noindent \textbf{Original Thresholds Discarded:} In PPF, the
perceptron sum is used to decide whether to prefetch or not, and the
fill-level in case of prefetch.  Thus, the confidence thresholds used 
by SPP -- T\textsubscript{f} and T\textsubscript{p} are no longer needed to
throttle the prefetcher and can be discarded.
%
%\pg{Note, however, as we discuss later, we
%use these confidence thresholds a feature in PPF so the mechanism is
%retained. -- right?}
% djimenez: keeping comments on every line here so it doesn't screw up my
%"roll-your-own" itemized list. don't put in any newlines.
%
%
% EB: Can this point be removed?
% djimenez: I think this would be necessary to reproduce our results so leave it in. 
% EB: It's ultimately a check to avoid SPP's internal segfault. Won't have a real impact.
%
%\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
%  original SPP does not consider the slots available in the L2 MSHR
%  queue.  It has the assumption that at no time can the number of
%  suggested prefetches exceed the capacity of the queue.  While it was
%  not described in the paper, the internal confidence mechanism makes
%  sure that the above assumption was maintained.  In PPF, since there
%  cannot be an assumption on the degree of aggressiveness of the
%  underlying prefetcher, we explicitly check that at no point should
%  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
%    this a big deal?  Seems like a detail}
%

@@

\subsection{Features used by Perceptron}
\label{Impl-Features}
Here we discuss the various features that correlate the prefetching
decision with the program behavior.  All the features we used can be
derived from the information available in the L2 Cache access stream
or are taken as metadata derived from the {\color{red}underlying} prefetcher.  Our
feature selection involved searching over a large space of relevant
perceptron features.  Note that part of the process of tuning PPF to a
specific prefetcher involves examining the available metadata in the
prefetcher itself, and thus PPF is attuned to the {\color{red}underlying}
prefetcher's design. Using the statistical methodology outlined in
Section~\ref{Method-Features}, we pruned the feature set to a minimal
yet relevant set of features.
% \pg{again you want to get at why
% these were chosen. Also mention that you searched over a large
% space of possible features and these were the best.}
%
% [EB] Tried addressing. Does it look redundant when read after 'Feature Selection'
% bullet from the previous section?

\noindent \textbf{Physical Address}: Here we use the lower bits 
of the physical address of 
the demand access that triggers the prefetch. This address corresponds 
to a  stream of accesses that SPP and the PPF have seen before. 
Therefore, PPF will correlate the past behavior of this address 
to the prefetch outcome. 
%
%\pg{what do you mean by ``base address'' the full physical
%  address?  Be specific.  Which bits and why? You should probably also
%  talk about what you do when the number of bits is unwieldy.}

\noindent \textbf{Cache Line} and \textbf{Page Address}: These two
separate features are derived from the base address that triggered the
prefetch, as follows: $base\_addr >> LOG2\_BLOCK \_SIZE$ and
$base\_addr >> LOG2\_PAGE\_SIZE$ respectively. The idea behind using
three different shifted versions of the same feature is that it allows
the filter to focus its examination in more detail on different
aspects of the address than with a single version.  It also helps give
more importance to the overlapping bits and lesser importance to most
and least significance bits.  This approach can also eliminate
destructive interference that can be caused by directly folding
the address bits into half.

%
% djimenez: this is a nice intuition and explanation. 
% [EB] Thank you!
\noindent \textbf{Program Counter XOR Depth}: The PC is for the
instruction that triggered the prefetch chain.  Depth refers to the
iteration count of the lookahead stages.  By itself, we find the PC to
not be a good basis for filtering a lookahead prefetcher, as all the
prefetches with depth >= 1 are aliased into the same PC, which will
not be the PC of the eventual actual demand access.  This feature
resolves a PC into a different value for each lookahead depth of
prefetch speculation, giving a more accurate correlation in lookahead
cases.  This is akin to the concept of Virtual Program
Counters~\cite{VPC} introduced by Kim \textit{et. al.} for indirect
branch prediction.

\noindent \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
  PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC
before the instruction that triggered the current prefetch.  Hashing
together the last three PCs informs PPF about the path that led to the
current demand access and helps capture and branching information of
the current basic block.  PCs are shifted in the increasing order of
history before being hashed together. This is done to avoid the
resultant value of zero when 2 or more PCs are the same. Additionally,
blurring the information as it gets older allows us to get a wider and
yet more approximate look into the program's history.

\noindent \textbf{Program Counter XOR Delta}: This feature tells us if
a given PC favors particular value(s) of delta.  As noted earlier,
while the PC alone does not convey useful information, this hash
resolves the PC into different values based on the tendency of that PC
to favor a certain delta.  Thus, the dynamic nature of different
instances of the same memory access instruction is captured here.

 % GAC: Seems more predictable than dynamic if its capturing 
 % what is essentially a streaming access pattern. Not quite sure how to rephrase,
 % if I'm correct
%
 % GAC: Again having trouble phrasing the below point. I was attempting to say
 % that despite SPP having low confidence due to speculation, PPF may 
 % reinforce prefetches with lower confidence if SPP has had a tendency to be 
 % correct in the past.
\noindent \textbf{Confidence}: The confidence, on a scale of 0 to 100,
used to throttle lookahead depth in the original SPP design.  While
the original confidence does not directly make the decision to
prefetch, PPF correlates it to the correctness of a proposed prefetch.
While the original SPP may have dismissed a prefetch due to running
further into speculation, PPF can use the original confidence as
indicator of not only when prefetches become less confident, but also
how likely a low confident speculation is correct in the context of
other features.

\noindent \textbf{Page Address XOR Confidence}: This feature scores
the tendency of each page to be prefetch friendly or prefetch averse.
It helps resolve a page into different entries depending on its
confidence for prefetching, which can vary during phases of a program
execution.

\noindent \textbf{Current Signature XOR Delta}: Recall from the
discussion of SPP in Section~\ref{Background-SPP} that the new
signature is generated using the old signature and the current block
delta.  The result of this feature is the next signature that is
predicted to be accessed based on the delta predicted by SPP. While
``Current Signature XOR Delta'' is not the exact formula for
generating the future signature, it gives an approximate idea of the
path that the combination of these two values can lead to.

%Gino: Added the sentence below, is this correct?
%[EB]: It gives the next signature instead of next delta
%
%%%%%%% Below are the features rejected when downsizing 14 -> 9
% \item \textbf{Current Signature}: The 12-bit signature used by SPP
%   to index into the Pattern Table
% \item \textbf{Delta}: The signed version of the difference in the
%   block address that triggers the prefetch and the block address of
%   the data being prefetched
% \item \textbf{PC}: Program Counter of the Load instruction that
%   triggers that particular prefetch
% \item \textbf{Base address XOR Delta}: If any particular value(s) of
%   delta are highly favored by a given address which triggered the
%   prefetch, this composite feature can capture that information
% \item \textbf{Cache line XOR Delta}: this feature captures if any
%   particular delta is highly favored by a given cache line address
%\item \textbf{Confidence XOR Delta}: This gives a combined insight
%  into that fact that certain delta-confidence pairs predicted by
%  SPP might correlate more with the prefetch outcome
%%%%%%%%%%%%%%%%%%%%%%

As can be noted above, some composite features are derived from simple
hashing (XOR) of two primary features. There is always a question of
usefulness of such composite features and the new information added.
We justify the choice of each feature by quantifying the contribution
made towards predicting prefetch behavior, in Section
\ref{Method-Features}.  Finally, as noted above, each feature indexes
into its independent entry of perceptron weights.

>>>>

\subsection{Changes Made to SPP}
\label{Impl-Changes}
\noindent To modify the SPP design to suit our scheme, the following changes
were made:
\newline
\newline
\noindent \textbf{Exporting Features from SPP:} PPF uses the metadata 
specific to SPP, to build some of the perceptron features. These include
the lookahead depth, signature and the confidence counter. These features 
were made visible to PPF.
%
\newline
\newline
\noindent \textbf{Original Thresholds Discarded:} In PPF, the
perceptron sum is used to decide whether to prefetch or not, and the
fill-level in case of prefetch.  Thus, the confidence thresholds used 
by SPP -- T\textsubscript{f} and T\textsubscript{p} are no longer needed to
throttle the prefetcher and can be discarded.
%
%\pg{Note, however, as we discuss later, we
%use these confidence thresholds a feature in PPF so the mechanism is
%retained. -- right?}
% djimenez: keeping comments on every line here so it doesn't screw up my
%"roll-your-own" itemized list. don't put in any newlines.
%
%
% EB: Can this point be removed?
% djimenez: I think this would be necessary to reproduce our results so leave it in. 
% EB: It's ultimately a check to avoid SPP's internal segfault. Won't have a real impact.
%
%\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
%  original SPP does not consider the slots available in the L2 MSHR
%  queue.  It has the assumption that at no time can the number of
%  suggested prefetches exceed the capacity of the queue.  While it was
%  not described in the paper, the internal confidence mechanism makes
%  sure that the above assumption was maintained.  In PPF, since there
%  cannot be an assumption on the degree of aggressiveness of the
%  underlying prefetcher, we explicitly check that at no point should
%  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
%    this a big deal?  Seems like a detail}
%

@@

\subsection{Features used by Perceptron}
\label{Impl-Features}
Here we discuss the various features that correlate the prefetching
decision with the program behavior.  All the features we used can be
derived from the information available in the L2 Cache access stream
or are taken as metadata derived from the {underlying} prefetcher.  Our
feature selection involved searching over a large space of relevant
perceptron features.  Note that part of the process of tuning PPF to a
specific prefetcher involves examining the available metadata in the
prefetcher itself, and thus PPF is attuned to the {underlying}
prefetcher's design. Using the statistical methodology outlined in
Section~\ref{Method-Features}, we pruned the feature set to a minimal
yet relevant set of features.
% \pg{again you want to get at why
% these were chosen. Also mention that you searched over a large
% space of possible features and these were the best.}
%
% [EB] Tried addressing. Does it look redundant when read after 'Feature Selection'
% bullet from the previous section?
%
\newline
\newline
\noindent \textbf{Physical Address}: Here we use the lower bits 
of the physical address of 
the demand access that triggers the prefetch. This address corresponds 
to a  stream of accesses that SPP and the PPF have seen before. 
Therefore, PPF will correlate the past behavior of this address 
to the prefetch outcome. 
%
%\pg{what do you mean by ``base address'' the full physical
%  address?  Be specific.  Which bits and why? You should probably also
%  talk about what you do when the number of bits is unwieldy.}
%
\newline
\newline
\noindent \textbf{Cache Line} and \textbf{Page Address}: These two
separate features are derived from the base address that triggered the
prefetch, as follows: $base\_addr >> LOG2\_BLOCK \_SIZE$ and
$base\_addr >> LOG2\_PAGE\_SIZE$ respectively. The idea behind using
three different shifted versions of the same feature is that it allows
the filter to focus its examination in more detail on different
aspects of the address than with a single version.  It also helps give
more importance to the overlapping bits and lesser importance to most
and least significance bits.  This approach can also eliminate
destructive interference that can be caused by directly folding
the address bits into half.
%
%
% djimenez: this is a nice intuition and explanation. 
% [EB] Thank you!
\newline
\newline
\noindent \textbf{Program Counter XOR Depth}: The PC is for the
instruction that triggered the prefetch chain.  Depth refers to the
iteration count of the lookahead stages.  By itself, we find the PC to
not be a good basis for filtering a lookahead prefetcher, as all the
prefetches with depth >= 1 are aliased into the same PC, which will
not be the PC of the eventual actual demand access.  This feature
resolves a PC into a different value for each lookahead depth of
prefetch speculation, giving a more accurate correlation in lookahead
cases.  This is akin to the concept of Virtual Program
Counters~\cite{VPC} introduced by Kim \textit{et. al.} for indirect
branch prediction.
%
\newline
\newline
\noindent \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
  PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC
before the instruction that triggered the current prefetch.  Hashing
together the last three PCs informs PPF about the path that led to the
current demand access and helps capture and branching information of
the current basic block.  PCs are shifted in the increasing order of
history before being hashed together. This is done to avoid the
resultant value of zero when 2 or more PCs are the same. Additionally,
blurring the information as it gets older allows us to get a wider and
yet more approximate look into the program's history.
%
\newline
\newline
\noindent \textbf{Program Counter XOR Delta}: This feature tells us if
a given PC favors particular value(s) of delta.  As noted earlier,
while the PC alone does not convey useful information, this hash
resolves the PC into different values based on the tendency of that PC
to favor a certain delta.  Thus, the dynamic nature of different
instances of the same memory access instruction is captured here.
%
 % GAC: Seems more predictable than dynamic if its capturing 
 % what is essentially a streaming access pattern. Not quite sure how to rephrase,
 % if I'm correct
%
 % GAC: Again having trouble phrasing the below point. I was attempting to say
 % that despite SPP having low confidence due to speculation, PPF may 
 % reinforce prefetches with lower confidence if SPP has had a tendency to be 
 % correct in the past.
\newline
\newline
\noindent \textbf{Confidence}: The confidence, on a scale of 0 to 100,
used to throttle lookahead depth in the original SPP design.  While
the original confidence does not directly make the decision to
prefetch, PPF correlates it to the correctness of a proposed prefetch.
While the original SPP may have dismissed a prefetch due to running
further into speculation, PPF can use the original confidence as
indicator of not only when prefetches become less confident, but also
how likely a low confident speculation is correct in the context of
other features.
%
\newline
\newline
\noindent \textbf{Page Address XOR Confidence}: This feature scores
the tendency of each page to be prefetch friendly or prefetch averse.
It helps resolve a page into different entries depending on its
confidence for prefetching, which can vary during phases of a program
execution.
%
\newline
\newline
\noindent \textbf{Current Signature XOR Delta}: Recall from the
discussion of SPP in Section~\ref{Background-SPP} that the new
signature is generated using the old signature and the current block
delta.  The result of this feature is the next signature that is
predicted to be accessed based on the delta predicted by SPP. While
``Current Signature XOR Delta'' is not the exact formula for
generating the future signature, it gives an approximate idea of the
path that the combination of these two values can lead to.
%
%Gino: Added the sentence below, is this correct?
%[EB]: It gives the next signature instead of next delta
%
%%%%%%% Below are the features rejected when downsizing 14 -> 9
% \item \textbf{Current Signature}: The 12-bit signature used by SPP
%   to index into the Pattern Table
% \item \textbf{Delta}: The signed version of the difference in the
%   block address that triggers the prefetch and the block address of
%   the data being prefetched
% \item \textbf{PC}: Program Counter of the Load instruction that
%   triggers that particular prefetch
% \item \textbf{Base address XOR Delta}: If any particular value(s) of
%   delta are highly favored by a given address which triggered the
%   prefetch, this composite feature can capture that information
% \item \textbf{Cache line XOR Delta}: this feature captures if any
%   particular delta is highly favored by a given cache line address
%\item \textbf{Confidence XOR Delta}: This gives a combined insight
%  into that fact that certain delta-confidence pairs predicted by
%  SPP might correlate more with the prefetch outcome
%%%%%%%%%%%%%%%%%%%%%%
%
\newline
\newline
As can be noted above, some composite features are derived from simple
hashing (XOR) of two primary features. There is always a question of
usefulness of such composite features and the new information added.
We justify the choice of each feature by quantifying the contribution
made towards predicting prefetch behavior, in Section
\ref{Method-Features}.  Finally, as noted above, each feature indexes
into its independent entry of perceptron weights.

### 2019-04-17 intro.tex
Paragraph spacing

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{Motivation}
\caption{The impact of aggressive prefetching on performance for {\tt 603.bwaves\_s}. 
The number of useful prefetches increases with aggressiveness
slower than total prefetches, which wastes bandwidth and 
harms performance.}
\label{Fig:Motivation}
\end{figure}

@@

Here, we propose Perceptron-based Prefetch Filtering (PPF) as an
enhancement to existing state-of-the-art prefetchers, allowing them to
speculate deeply to achieve high coverage while filtering out the
inaccurate prefetches this deep speculation implies.  PPF works by
observing the stream of candidate prefetches generated by a
prefetcher, and then rejects those that are predicted by the
online-trained neural model to be inaccurate.  The state-of-the-art
prefetcher that we use to evaluate PPF in this paper is the Signature
Path Prefetcher (SPP)~\cite{SPP}, however as we describe, PPF can be
designed to benefit any prefetcher.  In this design, PPF replaces
SPP's existing confidence-based throttling mechanism, which itself was
a highly tuned feature of that prefetcher.  Because PPF is so much
more effective at rejecting inaccurate prefetches than SPP's 
{\color{red}internal} mechanism, we are free to re-tune the rest of 
SPP's design around maximizing coverage. The result is an increase 
in both accuracy and coverage, and a notable increase in performance.

>>>>

\begin{figure}[t]
\includegraphics[width=0.95\columnwidth]{Motivation}
\caption{The impact of aggressive prefetching on performance for {\tt 603.bwaves\_s}. 
The number of useful prefetches increases with aggressiveness
slower than total prefetches, which wastes bandwidth and 
harms performance.}
\label{Fig:Motivation}
\end{figure}

@@

@@

Here, we propose Perceptron-based Prefetch Filtering (PPF) as an
enhancement to existing state-of-the-art prefetchers, allowing them to
speculate deeply to achieve high coverage while filtering out the
inaccurate prefetches this deep speculation implies.  PPF works by
observing the stream of candidate prefetches generated by a
prefetcher, and then rejects those that are predicted by the
online-trained neural model to be inaccurate.  The state-of-the-art
prefetcher that we use to evaluate PPF in this paper is the Signature
Path Prefetcher (SPP)~\cite{SPP}, however as we describe, PPF can be
designed to benefit any prefetcher.  In this design, PPF replaces
SPP's existing confidence-based throttling mechanism, which itself was
a highly tuned feature of that prefetcher.  Because PPF is so much
more effective at rejecting inaccurate prefetches than SPP's 
{internal} mechanism, we are free to re-tune the rest of 
SPP's design around maximizing coverage. The result is an increase 
in both accuracy and coverage, and a notable increase in performance.

### 2019-04-17 method.tex
Paragraph spacing

%This is the first time that SPEC CPU 2017 benchmark suite~\cite{SPEC2017}
%has been used to characterize and measure the prefetch performance.
%We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. 
We use all the 20 workloads available in the SPEC CPU 2017
suite~\cite{SPEC2017}. Using the SimPoint~\cite{SimPoint} methodology,
we identified 95 different program segments of 1 Billion
instructions each.

\noindent \textbf{Single-core performance:} For single-core
simulations, we use the first 200 million instructions to warm-up the
microarchitectural structures and the next one billion instructions to
do detailed simulations and collect run-time statistics. We report the
IPC speedup over the baseline of no prefetching.  The final numbers
reported are the geometric mean of the weighted mean speedup achieved
per application using the SimPoint methodology.

\noindent \textbf{Multi-core performance:} For multi-application
workloads, we generate 100 random mixes and another 100 mixes from the
memory intensive subset of SPEC CPU 2017.  For 4-core workloads, 200
Million instructions are used for warm-up and additional 1 Billion
instruction simulated for collecting statistics.  Each CPU keeps
executing its workload till the last CPU completes one billion
instructions after warm-up.  For collecting IPC and other data, only
the first billion instructions are considered as the region of
interest.

@@

We repeat the same process for 8-core workloads, correspondingly with
16MB LLC. The only difference is that 20 million warm-up instructions
and 100 million full instructions are executed. This is done so as to
keep the simulation run-time within reasonable limits as a single
8-core mix takes up to 3 days to simulate one billion instructions.

\noindent \textbf{Validation:} We cross-validated our PPF model using
SPEC CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite}
benchmarks. For single-core SPEC CPU 2006, we developed 94 simpoints
spread across all the 29 applications.  For multi-core, we followed
the same methodology as SPEC CPU 2017.  For CloudSuite, we used the
traces made available for the 2nd Cache Replacement Competition
(CRC-2)~\cite{CRC_2}. The traces include four 4-core applications with
six distinct phases per application.

@@

\subsection{Developing Features for PPF}
\label{Method-Features}
This section describes the intuition and analysis that went behind
developing the perceptron features.  As noted earlier, we developed a
set of nine features that allow the perceptron layer to correlate
prefetching decision with the program behavior.  To study the
correlation across each feature, we statistically examine the
perceptron weights and try to interpret their distribution.

\noindent \textbf{Global Pearson's Correlation:} Here we examine the
perceptron weights at the end of all trace execution by which time the
weights have settled to steady values. The weights obtained from
running all the SPEC CPU 2017 traces are concatenated.  Features with
a bulk of their perceptron weights concentrated around 0 or small
magnitude numbers show a weak correlation with the prefetching
outcome.  On the other hand, features with most of the weights
saturated around highest value (+15) show a high positive correlation
and the features with weights close to the lowest value (-16) show a
strong negative correlation.

@@

Figure~\ref{Fig:Pvals} shows all the features which are finally used,
arranged in the increasing order of their Pearson's factor.  As can be
seen 5 out of the 9 features provide a moderate to high correlation,
with the magnitude of P-value > 0.6. The single most important
feature, \textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90.

\noindent \textbf{Per Trace Correlation:} Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces. Here we give special attention to features
with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the variation
of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and \textit{PC XOR Depth}; across all the
SPEC CPU 2017 traces.  For simplicity, the traces are arranged in an
increasing order of contribution made by the feature.  It can be seen
that even features with a low overall correlation provide useful
correlation (magnitude > 0.5) for a significant number of traces. This
study motivated us to choose \textit{PC XOR DELTA} over \textit{Last
  Signature} as it provided useful correlation in at least some of the
traces.

\noindent \textbf{Trimming Features Using Cross-correlation:} Beside
providing interesting insights into prefetching behavior, P-value can
also be used for feature selection and prefetcher tuning.

@@

We initially came up {\color{red}with} a set of 23 features.  By studying cross
correlation of each of these features against others in a 23$\times$23
matrix, we identified pairs of features with correlation factor > 0.9
in magnitude and eliminated redundant features, using guidance from
Global and per-trace Pearson's factor of those features.  By doing
this, we reduced the feature count to 9.  Thus, in the final
implementation of PPF, no two features have a high correlation between
them. This way we can be sure that each feature makes a contribution
that cannot be captured using other features.

@@

In terms of computations, the perceptron mechanism only introduces an
extra adder tree. The hash perceptron mechanism makes sure than there
is no actual vector multiplication happening in the hardware.
Obtaining the perceptron sum requires addition of {\color{red}nine} 
5-bit numbers.
Using an adder tree of {\color{red}four} 5-bit adders, this can be done in
ceil($log_{2}9$) = 4 steps. Perceptron update only requires weight
update by +1 or -1. Thus, all the operations required for perceptron
inferencing or updating the states of the perceptrons can be easily
done in the time constraints of L2 Cache Accesses.

>>>>

%This is the first time that SPEC CPU 2017 benchmark suite~\cite{SPEC2017}
%has been used to characterize and measure the prefetch performance.
%We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. 
We use all the 20 workloads available in the SPEC CPU 2017
suite~\cite{SPEC2017}. Using the SimPoint~\cite{SimPoint} methodology,
we identified 95 different program segments of 1 Billion
instructions each.
\newline
\newline
\noindent \textbf{Single-core performance:} For single-core
simulations, we use the first 200 million instructions to warm-up the
microarchitectural structures and the next one billion instructions to
do detailed simulations and collect run-time statistics. We report the
IPC speedup over the baseline of no prefetching.  The final numbers
reported are the geometric mean of the weighted mean speedup achieved
per application using the SimPoint methodology.
\newline
\newline
\noindent \textbf{Multi-core performance:} For multi-application
workloads, we generate 100 random mixes and another 100 mixes from the
memory intensive subset of SPEC CPU 2017.  For 4-core workloads, 200
Million instructions are used for warm-up and additional 1 Billion
instruction simulated for collecting statistics.  Each CPU keeps
executing its workload till the last CPU completes one billion
instructions after warm-up.  For collecting IPC and other data, only
the first billion instructions are considered as the region of
interest.

@@

We repeat the same process for 8-core workloads, correspondingly with
16MB LLC. The only difference is that 20 million warm-up instructions
and 100 million full instructions are executed. This is done so as to
keep the simulation run-time within reasonable limits as a single
8-core mix takes up to 3 days to simulate one billion instructions.
\newline
\newline
\noindent \textbf{Validation:} We cross-validated our PPF model using
SPEC CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite}
benchmarks. For single-core SPEC CPU 2006, we developed 94 simpoints
spread across all the 29 applications.  For multi-core, we followed
the same methodology as SPEC CPU 2017.  For CloudSuite, we used the
traces made available for the 2nd Cache Replacement Competition
(CRC-2)~\cite{CRC_2}. The traces include four 4-core applications with
six distinct phases per application.

@@

\subsection{Developing Features for PPF}
\label{Method-Features}
This section describes the intuition and analysis that went behind
developing the perceptron features.  As noted earlier, we developed a
set of nine features that allow the perceptron layer to correlate
prefetching decision with the program behavior.  To study the
correlation across each feature, we statistically examine the
perceptron weights and try to interpret their distribution.
\newline
\newline
\noindent \textbf{Global Pearson's Correlation:} Here we examine the
perceptron weights at the end of all trace execution by which time the
weights have settled to steady values. The weights obtained from
running all the SPEC CPU 2017 traces are concatenated.  Features with
a bulk of their perceptron weights concentrated around 0 or small
magnitude numbers show a weak correlation with the prefetching
outcome.  On the other hand, features with most of the weights
saturated around highest value (+15) show a high positive correlation
and the features with weights close to the lowest value (-16) show a
strong negative correlation.

@@

Figure~\ref{Fig:Pvals} shows all the features which are finally used,
arranged in the increasing order of their Pearson's factor.  As can be
seen 5 out of the 9 features provide a moderate to high correlation,
with the magnitude of P-value > 0.6. The single most important
feature, \textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90. 
\newline
\newline
\noindent \textbf{Per Trace Correlation:} Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces. Here we give special attention to features
with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the variation
of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and \textit{PC XOR Depth}; across all the
SPEC CPU 2017 traces.  For simplicity, the traces are arranged in an
increasing order of contribution made by the feature.  It can be seen
that even features with a low overall correlation provide useful
correlation (magnitude > 0.5) for a significant number of traces. This
study motivated us to choose \textit{PC XOR DELTA} over \textit{Last
  Signature} as it provided useful correlation in at least some of the
traces.
\newline
\newline
\noindent \textbf{Trimming Features Using Cross-correlation:} Beside
providing interesting insights into prefetching behavior, P-value can
also be used for feature selection and prefetcher tuning.

@@

We initially came up {with} a set of 23 features.  By studying cross
correlation of each of these features against others in a 23$\times$23
matrix, we identified pairs of features with correlation factor > 0.9
in magnitude and eliminated redundant features, using guidance from
Global and per-trace Pearson's factor of those features.  By doing
this, we reduced the feature count to 9.  Thus, in the final
implementation of PPF, no two features have a high correlation between
them. This way we can be sure that each feature makes a contribution
that cannot be captured using other features.

@@

In terms of computations, the perceptron mechanism only introduces an
extra adder tree. The hash perceptron mechanism makes sure than there
is no actual vector multiplication happening in the hardware.
Obtaining the perceptron sum requires addition of {nine} 
5-bit numbers.
Using an adder tree of {four} 5-bit adders, this can be done in
ceil($log_{2}9$) = 4 steps. Perceptron update only requires weight
update by +1 or -1. Thus, all the operations required for perceptron
inferencing or updating the states of the perceptrons can be easily
done in the time constraints of L2 Cache Accesses.

### 2019-04-17 related.tex
Paragraph spacing

A low-accuracy aggressive prefetcher can significantly harm
performance.  To minimize interference from prefetching, Wu \textit{et
  al.} propose PACMan~\cite{pacman}, a prefetch-aware cache management
policy. PACMan dedicates some LLC sets to each of three competing
policies that treat demand and prefetch requests differently, using
the policy in the rest of the cache that shows the fewest
misses. Seshadri \textit{et al.} propose ICP~\cite{icp}, which demotes
a prefetch to the lowest reuse priority on a demand hit, based on the
observation that most prefetches are dead after their first hit. To
address prefetcher-caused cache pollution, it also uses a variation of
EAF~\cite{eaf} to track prefetching accuracy, and inserts only
accurate prefetches to the higher priority position in the LRU
stack. Jain \textit{et al.} propose Harmony~\cite{Harmony} to
accomodate prefetches in their MIN algorithm-inspired Hawkeye cache
management system.

@@

Peled \textit{et al.} introduce interesting ideas for on-line
Reinforcement Learning and dynamically scaling the magnitude of
feedback given to the baseline prefetcher~\cite{Semantics}. The
prefetcher relies on compiler support to receive features and build
the context.  Liao \textit{et al.}  focus on prefetching for data
center applications~\cite{Datacenter}.  They use offline machine
learning algorithms such as SVMs and logistic regression to do a
parametric search for an optimal prefetcher configuration.

>>>>

A low-accuracy aggressive prefetcher can significantly harm
performance.  To minimize interference from prefetching, Wu \textit{et
  al.} propose PACMan~\cite{pacman}, a prefetch-aware cache management
policy. PACMan dedicates some LLC sets to each of three competing
policies that treat demand and prefetch requests differently, using
the policy in the rest of the cache that shows the fewest
misses. Seshadri \textit{et al.} propose ICP~\cite{icp}, which demotes
a prefetch to the lowest reuse priority on a demand hit, based on the
observation that most prefetches are dead after their first hit. To
address prefetcher-caused cache pollution, it also uses a variation of
EAF~\cite{eaf} to track prefetching accuracy, and inserts only
accurate prefetches to the higher priority position in the LRU
stack. Jain \textit{et al.} propose Harmony~\cite{Harmony} to
accomodate prefetches in their MIN algorithm-inspired Hawkeye cache
management system. Ebrahimi \textit{et al.} introduce
HPAC~\cite{HPAC} which provides a coordinated control between 
multiple prefetchers present in a CMP by looking at the 
prefetcher-induced inter-core interferance.

@@

Peled \textit{et al.} introduce interesting ideas for on-line
Reinforcement Learning and dynamically scaling the magnitude of
feedback given to the baseline prefetcher~\cite{Semantics}. The
prefetcher relies on compiler support to receive features and build
the context.  Liao \textit{et al.}  focus on prefetching for data
center applications~\cite{Datacenter}.  They use offline machine
learning algorithms such as SVMs and logistic regression to do a
parametric search for an optimal prefetcher configuration. 
Hasheni \textit{et al.}~\cite{LSTM} categorize 
prefetching as a regression problem and use LSTM based Deep 
Learning approach.
